---
title: "STAT 440/840 - CM 761 - Assignment 2 - Spring 2019"
author: "Name"
date: 'Due: Tuesday, June 25 at 9:00am on Crowdmark'
output:
  pdf_document: default
  html_document: default
---


- **Assignment format policy**: Using RMarkdown or LaTeX is required and no hand written and/or imported screenshots will be accepted in the assignments. A mark of 0\% will be assigned to the questions which were not complied in RMarkdown or LaTeX, and/or those which include hand-written solutions and/or screenshots.

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```


1. The Weibull distribution with shape parameter $\alpha >0$ and scale parameter $\sigma>0$ is given by

$$ f(x \left|  \sigma,\alpha) \right. =
\begin{cases}
\frac{\alpha}{\sigma}\left(\frac{x}{\sigma}\right)^{\alpha-1}e^{-(x/\sigma)^{\alpha}} & x\geq0 ,\\
0 & x<0.
\end{cases}
$$
The MLE for the $\sigma$ given $\alpha$ is
$$ \widehat{\sigma} = \left[ \frac{\sum_{i=1}^n x_i^{\alpha} }{n} \right]^{1/\alpha} $$


- Assuming $\alpha=5.2$ use the following methods to construct confidence interval for $\sigma$
    i) standard bootstrap confidence interval (assuming normality),
    ii) bootstrap percentile interval,  
    iii) bootstrap-$t$ interval using the observed information, 
    iv) bootstrap-$t$ interval using the double bootstrap, 
    v) the asymptotic normality of the MLE and 
    vi) the log-likelihood ratio statistic. 
    
- **Note:**
    - For each method that requires bootstrapping, sample from the data with replacement.    
    - A clear answer with appropriate conclusions will obtain full marks.

$\;$

a) **[8 Marks]** Construct confidence interval for $\sigma$ using the following data and each of the method given above.

```{r}
temp = read.csv( "eng-monthly-011942-112007.csv" )
speed = temp$Spd.of.Max.Gust..km.h.
```

b) **[12 Marks]** Using a simulation study with  $(\sigma, \alpha)=(64,5.2)$ and $m=1000$ datasets, compare the coverage probabilities of the above confidence intervals. 


$\;$

$\;$

$\;$



2. In this question you will derive and implement an EM algorithm to fit a multivariate-normal distribution to Ozone $(z)$ and Wind $(x)$ from the air quality dataset. Make sure you define any notation that you introduce. 

```{r}
data(airquality)
head(airquality)
```

a) **[1 Mark]** What is the joint distribution of the missing data and the observed data? 

b) **[2 Marks]** State the conditional distribution of the missing data given the observed data. 

c) **[2 Marks]** State the complete data likelihood 

d) **[4 Marks]** E-step: Derive the expected complete data log-likelihood. 

e) **[4 Marks]** M-step: Derive the updates for the parameters. 

f) **[6 Marks]** Implement the above EM in `R` for the air quality dataset. Use starting values based on the complete cases. Give the MLE and plot the observed log-likelihood evaluated at each iteration. 

g) **[1 Mark]** Plot the imputed dataset.



